---
title: "YARGODYNE Corporation - NLP Algo"
author: "Alejandro Borges Sanchez"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## YARGODYNE NLP Research Task Force Targets

As tasked, the team has developed infrastructure for the generation and evaluation of NLP models:

-   Latest stand-alone, high compression, high performance NGram algorithm called Knesser-Ney smoothing, as implemented by [Keneth Heafield](https://kheafield.com/code/kenlm/).

-   This allowed us to choose the smallest possible yet performant set of n-grams to run in our constrained production environment.

-   Considerable time was spent in building a thoroughly comprehensive and scalable analysis pipeline, this will enable us to update our model as the app is used.

## Slide with Bullets

-   Bullet 1
-   Bullet 2
-   Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```
